Does It Scale? Symbolic Execution of PHP Web Applications

### 2017-01-21 paper.tex
related work finished, conclusion draft

\definecolor{lightgrey}{HTML}{FFFFFF}

@@

\begin{document}
\title{Does It Scale? Symbolic Execution of PHP Web Applications}
%\titlenote{Produces the permission block, and copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as  \texttt{acmart.pdf} document}

@@

One common property of all technologies for dynamic web applications is
\emph{staged computation}: A dynamic web application as a whole
consists of both static code, such as scripts, and dynamically generated code,
such as client page output. The latter code, although assembled at
runtime, may contain client-side parts of the web application, such as
JavaScript. So, to study dynamic web applications in its entirety, we need to
consider both static as well as dynamic aspects of systems.

@@

Second, we ask :Do system size and usage of modern language features
factors affect the practicality of our symbolic execution engine? We
selected both a range of large-scale and small-scale modern PHP systems to
better separate and understand the circumstances under which conceptual
limitations might affect practicality. For the selection of large-scale systems
we borrow three systems from a case study addressing the feature usage in PHP
systems. Some more detailed description of this case study can be found in
Section \ref{sec:related_work}. For the small-scale systems we selected five systems from
a list of recent Content Management Systems (CMS) written in PHP \cite{codegeekz}. For both
selections we were bound to select a smaller number of system than we intended
to since the parser used in our implementation, \sf{Quercus}\footnote{Quercus
footnote}, did not support a number of language features. The full list of PHP
systems for that we evaluate our symbolic execution engine along with
descriptive code metrics can be found in Table \ref{corpus}.

@@

% RESULT DIAGRAMS
\begin{figure}[h!]
	\begin{subfigure}[center]{0.48\textwidth}
		\input{new_diagrams/not_reached}
		\caption{\label{coverage}}
	\end{subfigure}

	\begin{subfigure}[center]{0.48\textwidth}
		\input{new_diagrams/includes}
		\caption{
			Include coverage and accuracy.
			\label{fig:include_coverage_results}
		}
	\end{subfigure}

	% Callback candidate explanation
	\begin{subfigure}[center]{0.48\textwidth}
		\input{new_diagrams/callbacks}
		\caption{
			Distribution dead function output candidates that are par-
			tially or completely explained due to callback candidates.
			\label{fig:output_candidate_explanation}
		}

	\end{subfigure}
	\caption{Results}
\end{figure}

\subsubsection{What literals did we miss?}\label{WhatLiteralsDidWeMiss?}
Our initial approach to understand missed output candidates is to find out
whether their surrounding program code was not accessed/accessible, and if so,
why. We are interested in those output candidates that are located in HTML
files (possibly with nested PHP), or located in a function definition. Plain
HTML files do not require any further execution and represent output candidates
that just need to be included properly to be part of the output. Aside, output
candidates that are part of a function definition are only missed if the
corresponding function is never called at runtime.

@@

\subsubsection{Why did includes fail?}
\label{WhyDidIncludesFail}
From Section \ref{sec:experiment_dynamicfeatures} we learn that include
expressions can not be evaluated accurately for symbolic values, and from the
previous Section \ref{WhatLiteralsDidWeMiss?} that very little of missed output
candidates can be explained by simply missing to include a file. To better
understand the impact of include expression resolution throughout analyzing
systems, we construct two additional metrics. First, we are interested in how
many include expressions throughout the execution we actually reach in a system.
Therefore we introduce the metric \emph{reach coverage for include expressions}
as the ratio of reached include expressions and the total number of reach
expressions in a system to check whether failed includes may have caused
subsequent include expressions and so on. Second, we measure how successful
include expressions are resolved.

@@

\subsubsection{Why did function calls fail?}
\label{WhyDidFunctionCallsFail}
Aside from failed or successful includes, this metric describes one scenario of
why functions containing not-reached output candidates are never called: The
trivial explanation would be that there are no direct calls of those functions,
yet PHP offers a number of ways to call a function indirectly, for instance
using built-in functions, or simply evaluating strings as PHP code using the
\texttt{eval} function.

@@

\paragraph{Bypassing conceptual limitations.}
Symbolic execution is used to approximate client page output since it allows to
unfold the staged nature of dynamic web applications. Aside from conceptual
limitations in the systems studied, limitations can be addressed. To use
approaches incorporating symbolic execution with dynamic web applications in
the long run, more concrete dummy information, similar to test oracles, needs
to be provided to decrease the number of assumed symbolic inputs and avoid
dynamic features being confronted with symbolic values. Also, if the usage of
tools using static output approximation with symbolic execution is intended,
avoiding dynamic features that are prone to be inaccurately executed for
symbolic execution  may be an adjustment strategy to program in an analyzable
way. 
Future work to increase code coverage and approximation quality might involve
automated test case generation as tools like \sf{Apollo}
\cite{artzi_finding_2008,artzi_finding_2010} or previous
work \cite{DynamicWassermann} to decrease the number of symbolic inputs and
provide proper and concrete input variants.

@@

\section{Related Work} \label{sec:related_work}
\paragraph{Static Output Approximation}
To approximate program slices for web
applications, Ricca et al. \cite{tonella_web_2005,tonella_2001,tonella_2002} approximate
dynamically generated output. For output generating statements, such as \tt{echo} or
\tt{print}, all strings are unquoted (code extrusion). If those statements
contain variables, these are linked to string concatenations using a
proposed flow analysis called \emph{string-cat propagation}. From these
flows representing approximated output subsequent program slices are computed.

@@

\paragraph{Input Generation for PHP}
\textcolor{gray}{Apollo \cite{artzi_finding_2008,artzi_finding_2010}, automated
test input generation using concolic testing \cite{DynamicWassermann}, test case generation
using search based software engineering \cite{Alshahwan2011}}

@@

\section{Conclusion}

>>>>

\definecolor{lightgrey}{HTML}{FCFCFF}

@@

\begin{document}
\title{Does It Scale? Symbolic Execution of PHP Web Applications}
%\titlenote{Produces the permission block, and copyright information}
\subtitle{Output-oriented Symbolic Execution for Static Output Approximation}
%\subtitlenote{The full version of the author's guide is available as  \texttt{acmart.pdf} document}

@@

A common characteristic of all dynamic web applications is their
\emph{staged nature}: A dynamic web application as a whole
consists of both static code, such as scripts, and dynamically generated code,
such as client page output. The latter code, although assembled at
runtime, may contain client-side parts of the web application, such as
JavaScript. So, to study dynamic web applications in its entirety, we need to
consider both static as well as dynamic aspects of systems.

@@

Second, we ask: Do system size and usage of modern language features
factors affect the practicality of our symbolic execution engine? We
selected both a range of large-scale and small-scale modern PHP systems to
better separate and understand the circumstances under which conceptual
limitations might affect practicality. For the selection of large-scale systems
we borrow three systems from a case study addressing the feature usage in PHP
systems. Some more detailed description of this case study can be found in
Section \ref{sec:related_work}. For the small-scale systems we selected five systems from
a list of recent Content Management Systems (CMS) written in PHP \cite{codegeekz}. For both
selections we were bound to select a smaller number of system than we intended
to since the parser used in our implementation,
\sf{Quercus}\footnote{\sf{Quercus} is an open-source Java implementation of PHP
5 by Caucho Technology. We reused the parser front-end for our implementation.
\sf{Quercus} can be obtained from \url{http://quercus.caucho.com}.}, did not
support a number of language features.
The full list of PHP systems for that we evaluate our symbolic execution engine along with descriptive code metrics can be found in Table \ref{corpus}.

@@

% RESULT DIAGRAMS
\begin{figure}[h!]
	\begin{subfigure}[center]{0.48\textwidth}
		\input{new_diagrams/not_reached}
		\caption{
		Classification distribution of \emph{reached} output candidates
		(Reached/Output, OnlyReached) and \emph{not-reached} output candidates
		(FunctionContext, NotImported, Misc) during the analyis for each system.
		\label{coverage}}
	\end{subfigure}
	\vspace{1mm}

	\begin{subfigure}[center]{0.48\textwidth}
		\input{new_diagrams/includes}
		\caption{
			Classification distribution of \emph{reached} output include expressions
			(Reached/Output, OnlyReached) and \emph{not-reached} include expressions.
			\label{fig:include_coverage_results}
		}
	\end{subfigure}
	\vspace{1mm}

	% Callback candidate explanation
	\begin{subfigure}[center]{0.48\textwidth}
		\input{new_diagrams/callbacks}
		\caption{
			Distribution of \emph{not-reached} output candidates (in function contexts)
			that are only (Complete), partially (Partial), or not accessible via callback candidates. 
			\label{fig:output_candidate_explanation}
		}

	\end{subfigure}
	\caption{Evaluation results for output candidates, include expressions and
	callgraph analysis for not-reached output candidates}
\end{figure}

\subsubsection{Where did we miss output?}\label{WhatLiteralsDidWeMiss?}
Our initial approach to understand missed output candidates is to find out
whether their surrounding program code was not accessed/accessible, and if so,
why. We are interested in those output candidates that are located in HTML
files (possibly with nested PHP), or located in a function definition. Plain
HTML files do not require any further execution and represent output candidates
that just need to be included properly to be part of the output. Aside, output
candidates that are part of a function definition are only missed if the
corresponding function is never called at runtime.

@@

\subsubsection{(Dynamic) Includes}
\label{WhyDidIncludesFail}
From Section \ref{sec:experiment_dynamicfeatures} we learn that include
expressions can not be evaluated accurately for symbolic values, and from the
previous Section \ref{WhatLiteralsDidWeMiss?} that very little of missed output
candidates can be explained by simply missing to include a file. To better
understand the impact of include expression resolution throughout analyzing
systems, we construct two additional metrics. First, we are interested in how
many include expressions throughout the execution we actually reach in a system.
Therefore we introduce the metric \emph{reach coverage for include expressions}
as the ratio of reached include expressions and the total number of reach
expressions in a system to check whether failed includes may have caused
subsequent include expressions and so on. Second, we measure how successful
include expressions are resolved.

@@

\subsubsection{Dynamic Function Invocations}
\label{WhyDidFunctionCallsFail}
Aside from failed or successful includes, this metric describes one scenario of
why functions containing not-reached output candidates are never called: The
trivial explanation would be that there are no direct calls of those functions,
yet PHP offers a number of ways to call a function indirectly, for instance
using built-in functions, or simply evaluating strings as PHP code using the
\texttt{eval} function.

@@

\paragraph{Bypassing conceptual limitations.}
Symbolic execution is used to approximate client page output since it allows to
unfold the staged nature of dynamic web applications. Aside from conceptual
limitations in the systems studied, limitations can be addressed. To use
approaches incorporating symbolic execution with dynamic web applications in
the long run, more concrete input, similar to test oracles, needs
to be provided to decrease the number of assumed symbolic inputs and avoid
dynamic features being confronted with symbolic values. Also, if the usage of
tools using static output approximation with symbolic execution is intended,
avoiding dynamic features that are prone to be inaccurately executed for
symbolic execution  may be an adjustment strategy to program in an analyzable
way.

@@

\section{Related Work} \label{sec:related_work}

\paragraph{Symbolic Execution}
Symbolic execution \cite{King1976,Darringer1978} remained an conceptual approach
for decades, but has been refined and extended over the last years. Several
approaches incorporating constraint solving and leveraging concrete execution
anlongside symbolic execution emerged and enable automated test case generation
and testing; the scalability of these approaches is still challenged by path
explosion and constraint solving \cite{CadarSen2013}.

Dynamic symbolic execution \cite{CadarSen2013} incorporates symbolic execution
alongside concrete execution and ahs shown promising results in automated test
generation and testing \cite{artzi_finding_2008,artzi_finding_2010,DynamicWassermann} for PHP web
applications and might help mitigate the conceptual limitations faced with
symbolic values and dynamic features for future work on output-oriented
symbolic execution.

An overview about recent evolution of symbolic execution is provided by work of Cadar et al.
\cite{CadarSen2013}, as well as a more tool-centered perspective \cite{Cadar2011}.

\paragraph{Static Output Approximation}
To approximate program slices for web
applications, Ricca et al. \cite{tonella_web_2005,tonella_2001,tonella_2002} approximate
dynamically generated output. For output generating statements, such as \tt{echo} or
\tt{print}, all strings are unquoted (code extrusion). If those statements
contain variables, these are linked to string concatenations using a
proposed flow analysis called \emph{string-cat propagation}. From these
flows representing approximated output subsequent program slices are computed.

@@

@@

\section{Conclusion}
In this paper we have addressed symbolic execution of PHP web applications that
is used for client page output approximation. This approximations are leveraged
be several developer tools for PHP, yet have not been evaluated for larger
systems. We re-implemented and extended an existing symbolic execution engine
to analyze several real-world PHP applications with the intention to evaluate
the scalability of symbolic execution for PHP. Our work illustrates the
necessity of more comprehensive evaluation of developer tools as
previous work did not explore these limitations. Dynamic features of PHP in
combination with symbolic values though hinder precise symbolic execution and
limit the scalability and practicality of tools on symbolically approximated
client page output.

In the future we would like to use dynamic symbolic execution for output
approxiamtion and possibly bypass existing limitations, as dynamic symbolic
execution has shown to be promising for automated testing and verification.
